---
title: "hw3_nn2477"
author: "Nhu Nguyen"
date: "2023-10-11"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
## problem 1
```{r}
data("instacart")
instacart %>% 
  summarise(count = n_distinct(aisle))
```

## getting the aisles with the most item ordered
```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarise(total_items = n()) %>% 
  arrange(desc(total_items))
```

## making plot for aisle and number of items ordered 

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarise(total_items = n()) %>% 
  filter(total_items > 10000) %>% 
  arrange(desc(total_items)) %>% 
ggplot(aes(x = reorder(aisle, -total_items), y = total_items)) +
  geom_point() + 
  labs (
    title = "number of items ordered in each aisle", 
    x = "aisle",
    y = "total items ordered"
  ) + 
  coord_flip() + 
  theme_minimal()
```

## making a table for the three most popular items 

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(order_count = n()) %>% 
  arrange(aisle, desc(order_count)) %>% 
  group_by(aisle) %>% 
  slice_max(order_count, n = 3) %>% 
  pivot_wider(
    names_from = aisle,
    values_from = order_count)
```

## create table for pink landy apples and coffee ice cream 
```{r}
instacart %>% 
  mutate(order_dow = recode(
    order_dow,
    "0" = "sunday",
    "1" = "monday",
    "2" = "tuesday",
    "3" = "wednesday",
    "4" = "thursday",
    "5" = "friday",
    "6" = "saturday")) %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(order_dow, product_name) %>% 
  summarise(mean_hour = mean(order_hour_of_day, na.rm = TRUE)) 
```

## problem 2 
loading dataset 
```{r}
data("brfss_smart2010")
```

## cleaning dataset 

```{r}
brfss_df <- brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(county = locationdesc, state = locationabbr) %>% 
  filter(topic == "Overall Health",
     response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  mutate(response = factor(
    response,
    levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), exclude = NULL),
    county = gsub("-", "", as.character(county)),
    county = substring(county, first = 4))

```
## 2002 data

```{r}
brfss_2002 <- brfss_df %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarise(locations = n()) %>% 
  filter(locations >= 7) 
```
In 2002,`r with(brfss_2002, paste(unique(state), collapse = ", "))` were observed to have 7 or more locations. 

## 2010 data
```{r}
brfss_2010 <- brfss_df %>% 
  filter(year == 2010) %>% 
  group_by(state) %>% 
  summarise(locations = n()) %>% 
  filter(locations >= 7) 
```
In 2010,`r with(brfss_2010, paste(unique(state), collapse = ", "))` were observed to have 7 or more locations. 

```{r}
brfss_excellent <- brfss_df %>% 
  filter(response == "Excellent") %>% 
  group_by(year, state) %>% 
  summarise (mean_data_value = mean(data_value))
```
This data set has `r nrow(brfss_excellent)` observations that responded "Excellent" when asked about their general health

```{r}
ggplot(brfss_excellent, aes(x = year, y = mean_data_value, group = state, color = state)) +
  geom_line(alpha = .5) +
  labs(
    title = "average value over time by state",
    x = "year",
    y = "average data value (%)"
  ) + 
    theme_minimal()
```

this "spaghetti" plot shows a general trend of data value points decreasing over time for each state. the average data value ranges from 17.5 to 27.5 capturing most points, however some average data values do fall outside this range. specifically, WV appears to be a visual outlier in this plot. 

## making two panel plot for NYC data

```{r}
brfss_ny_df = brfss_df %>% 
  filter(year %in% c("2006", "2010"),
         state == "NY",
         response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) 

brfss_ny_df %>% 
  ggplot(aes(x = response, y = data_value)) + 
  geom_boxplot() +
  geom_point(aes(color = county)) +
  facet_grid(. ~ year) +
  labs(
    title = "distribution of data value by response in NY state counties in 2006 and 2010",
    x = "response", 
    y = "data_value (%)",
    color = "county"
  ) + 
  theme_minimal()
```

For both 2006 and 2010, the response "good" and "very good" has the highest data values while the response "poor" had the lowest data value. 

## problem 3 

cleaning, tidying, merging datasets 
```{r}
accel_df =
  read.csv("./data/nhanes_accel.csv") %>% 
  janitor::clean_names() 

covar_df = 
  read_csv("./data/nhanes_covar.csv", skip = 4) %>%
  janitor::clean_names() %>% 
  filter(age >= 21) %>% 
  mutate(sex = recode(
    sex, `1` = "male", `2` = "female"),
    education = recode(
      education, `1` = "less than high school", `2` = "high school equivalent", `3` = "more than high school"
    ),
    sex = factor(sex), 
    education = factor(education)) %>% 
  drop_na()

mims_df = left_join(covar_df, accel_df, by = "seqn")

```
men/women by education and age distribution
```{r}
education_mims_df <- mims_df %>% 
  group_by (sex, education) %>% 
  summarise(count = n()) %>% 
  pivot_wider(
    names_from = sex,
    values_from = count)
knitr::kable(education_mims_df, caption = "# of men and women in each education category")
```

the highest education both genders received was in the "more than high school" category.

```{r}
ggplot(mims_df, aes(x = education, y = age, fill = sex)) + 
  geom_boxplot(alpha = .5) + 
  labs(
    title = "age distribution by education and gender",
    x = "education category",
    y = "age"
  ) + theme_minimal()
```

females, on average, appeared to be older in the "high school equivalent" category, whereas males, on average, appeared to be older in the "less than high school" and "more than high school" education category. 

## total activity over day plot 

aggregating min variable
```{r}
total_activity_df <- mims_df %>% 
  mutate(
    total_activity = rowSums(across(c(min1:min1440)))
  )
```

creating plots:
```{r}
ggplot(total_activity_df, aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = .5) + 
  facet_wrap(~education, ncol = 1) + 
  geom_smooth(se = FALSE)+
  labs(
    title = "total activity vs age by education level", 
    x = "age",
    y = "total activity"
  ) + 
  theme_minimal()
  
```

in all 3 education categories, on average, the total physical activity for both genders tend to decrease with overall age. for both "high school equivalent" and "more than high school" education category, women tend to be more active compared to their male counterparts (as seen in the trend lines). the greatest decline in trend line for both genders is in the "less than high school" education category.

## activity over day plot
```{r}
time_course_data <- total_activity_df %>% 
  group_by(education, sex) %>% 
  summarise(across(starts_with("min"), ~mean(.), .names = "mean_{.col}")) %>% 
  pivot_longer(
    cols = starts_with("mean"),
    names_to = "time",
    values_to = "mean") %>% 
  mutate(time = substring(time, 9),
         time = as.numeric(time))

ggplot(time_course_data, aes(x = time, y = ))
```


